\documentclass[11pt,onecolum,a4paper,oneside]{book}
\usepackage{a4}
\usepackage{graphicx,epsfig,subfigure}
\usepackage{amssymb}
\usepackage{ifthen}
\usepackage{pseudocode}
\usepackage{doublespace}
\usepackage{anuthesis}

\title{Hybrid Topological-Metric Simultaneous Localisation and
  Mapping} 
\author{Kirill Kouzoubov}

%Style 
%% \setlength\textwidth{171mm}
%% \setlength\oddsidemargin{-6.4mm}    %25.4 - 6.4 = 19
%% \setlength\evensidemargin{-6.4mm}   %25.4 - 6.4 = 19
%% \setlength\textheight{244mm}
%% \setlength\topmargin{-15.4mm}        %25.4 - 5.4 = 20


\begin{document}
\date{30 June 2006}

\maketitle
\normalsize

\onehalfspace
\chapter*{Declaration}

This thesis is an account of research undertaken between March 2002
and June 2006 at The Department of Systems Engineering, Faculty of
Engineering, The Australian National University, Canberra, Australia.

Except where acknowledged in the customary manner, the material 
presented in this thesis is, to the best of my knowledge, original and 
has not been submitted in whole or part for a degree in any 
university.

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank my supervisors Dr. David Austin and Prof. John
Moore. Special thanks to Dr. Alex Zelinsky for encouraging me to take
up the scholarship while I was a summer scholar in RSISE.

I express my gratitude to the organisers and hosts of the SLAM summer
school of 2002. I have learned a lot about SLAM over that week.
Presentations and code examples were extremely helpful.  Sandwiches
with cheese and capsicum slices, served in the afternoon, were of
great help too.

I have visited Intelligent Robot Laboratory, University of Tsukuba in
the winter of 2002. I have enjoyed three months in Japan, even though
it was rather cold sometimes. I would like to thank Prof. Shinichi
Yuta and Dr. Masahiro Tomono for the supervision during that time and
I thank the team from Roboken for their hospitality and sake.

%English
Dr. David Leibowitz and Dr. Nick Barnes have helped with English in
most parts of this thesis, readability has improved significantly
thanks to their efforts. Thank you to Rosemary Shephered for her great
help with the administrative side of this thesis.

%People from RSISE


%Vlada
Finally I acknowledge the contribution my wife Vlada has made to this
thesis, her support during the final stages of this thesis, were
crucial to the successful completion.





\input{notation_definitions.tex}

\chapter*{Abstract}  % the * means don't put a number in the title
\addcontentsline{toc}{chapter}{Abstract} % but add it to the table of contents

%\begin{abstract}
This thesis deals with the problem of simultaneous localisation and
mapping (SLAM). The problem of SLAM has been studied over the past two
decades by many researchers. Limitations of traditional approaches are
well understood in the research community. Some of the more
challenging aspects of the SLAM problem include: ambiguous data
association in the presence of clutter, loop closing, large maps,
dynamic environments.

Traditional approaches to SLAM cannot operate in large
environments. Computational complexity is one of the limiting
factors. A more important limitation is the accumulation of systematic
approximation errors that ultimately lead to an inconsistent map. It is
this author's belief that new map representations are needed to
overcome these limitations.

This thesis presents a novel map representation that combines metric
and topological information - hybrid topological SLAM (HTSLAM). The
map is a graph: local maps form nodes of the graph, each local map has
its own reference frame and edges of the graph maintain spatial
information and coordinate transformations between adjacent
regions. There is no global reference frame. FastSLAM is used for
building local maps. The uncertainty in the coordinate transformations
between adjacent maps is modelled using particles. This representation
allows constant time map updates and robust loop closing. Since there
is no global reference frame, this approach does not suffer from
constantly increasing uncertainty.

The ideas of this thesis are tested on real-life data. Three different
data sets are considered: a laser range scanner detecting corners in
indoor environment, a laser range scanner detecting trees in the park
(the well known Victoria Park data set), and a vision system detecting
vertical edges in an indoor environment.


%\end{abstract}

\tableofcontents
\listoffigures


\chapter{Literature review}
\input{chpt_LiteratureReview.tex}


\chapter{Background Theory and FastSLAM}
\label{chpt:Overview}
\input{chpt_FastSLAM.tex}


\chapter{HTSLAM Map Structure}
\label{chpt:MapStructure}
\input{chpt_MapStructure.tex}

\chapter{Mapping Procedure}
\label{chpt:Mapping}
\input{chpt_Mapping.tex}

\chapter{Multiple Hypothesis and Loop Closing}
\label{chpt:LoopClosing}
\input{chpt_LoopClosing.tex}

\chapter{Summary and Comparison}
\input{chpt_Summary.tex}

\chapter{Experimental Results: Laser Scanner}
\label{chpt:Laser_Results}
\input{chpt_ExperimentLaser.tex}

\chapter{Experimental Results: Vision}
\label{chpt:Vision_Results}
\input{chpt_ExperimentVision.tex}

\chapter{Conclusion}
\input{chpt_Conclusion.tex}

\bibliographystyle{plain}
\bibliography{my_ref,bio_ref,ref_slam}

\end{document}

% LocalWords:  HTSLAM FastSLAM Zelinsky RSISE Tsukuba Shinichi Yuta Masahiro
% LocalWords:  Tomono Roboken Vlada
